On the properties of discs around accreting brown dwarfs

### 2010-05-18 BD_disc/paper.tex
changes to intro - th
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@2269 25464f21-6099-402c-b30a-640f1d40d38b

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2009}

@@

Theoretical study of Brown Dwarfs (BDs) and the subsequent development
of testable hypotheses for observational studies is a critical, and
somewhat uncertain, area within the field of star and planet
formation. Several of the key physical parameters and mechanisms
remain unclear within this mass regime \citep{2007ApJ...664.1154S}.
Uncertainty remains over fundamental issues such as the initial
formation mechanisms, the subsequent evolution and dominance of a
star-disc interaction through to the possible dependence of planet
formation mechanisms on stellar mass \citep[for a review of BD
formation mechanisms see][]{2006astro.ph..2367W}.

Perhaps the two most fundamental challenges are formulating the
formation mechanism of BDs and deriving the resulting
initial-mass-function (IMF). Whether the BD IMF or formation
mechanisms are different to their higher mass, solar-type, equivalents
(T Tauri Stars, hereafter termed TTS) is an open question. Derivations
of reliable IMFs depend, obviously, on robust derivations of stellar
mass, and more subtly, on derivations of stellar ages (as IMFs are
usually defined for assumed coeval populations). The study of
classical T Tauri (CTTS) and disc systems of
\cite{1997AJ....114..288M} found that given an expected range of
parameters the scatter in infrared photometry matched the observed
CTTS locus well. However the work of \cite{1997AJ....114..288M},
discussed in more detail later, only included parametrised `flat'
discs. Once disc flaring is considered it is clear from previous work
that misidentification of high mass stars with circumstellar discs is
likely. \cite{2004ApJ...617.1177W} show that high-mass (HM) stars with
circumstellar discs are likely to be mis-identified as younger Class I
sources. This problem also exists at lower masses where the presence
of circumstellar discs can dramatically shift the spectral energy
distributions (SEDs) and therefore colours and magnitudes of BDs
\citep{2004MNRAS.351..607W}. Therefore, deriving robust masses (and
ages), and subsequently IMFs for BDs requires an understanding of
these effects of disc presence.

The presence, or fraction (for populations) of associated discs is
itself also crucial in isolating the formation mechanisms of
BDs. Precise determination of disc parameters for BDs and comparison
with the TTS results may allow the differentiation of currently
competing theories of the BD formation mechanism. Essentially, a
mechanism similar to that seen in higher mass stars i.e. cloud
collapse and fragmentation will result in a similar fraction of discs
across the BD mass boundary, whereas a formation mechanism dominated
by dynamical interactions \citep[ejections, accretion disc stripping
from near collisions or tidal interactions etc.]{2003IAUS..211...27B}
will result in a markedly different disc parameters, such as fractions
of stars with discs or radial sizes of the discs themselves, across
the BD mass boundary. Fundamentally a different disc fraction for BD
stars to TTS would provide persuasive evidence of a different
formation mechanism across the two mass regimes, as would an increase
in the average radial size of discs, indicative of truncation for BD
objects. Therefore, to progress our understanding of the fundamental
issues of both the formation mechanism and emerging IMFs for BDs,
consideration of the effects of an associated circumstellar disc and
their subsequent ubiquity are essential.

Another fundamental, and important, unresolved issue is the mechanism
of planet formation. A critical, and currently ill defined constraint
for the planet formation models is the timescale available for planet
formation within the circumstellar disc. Current favoured formation
mechanisms such as core accretion or gravitational instability are
separated by these timescales. Disc fractions have again been used, in
conjunction with stellar ages, to derive the disc dissipation
timescale, which obviously constrains the planet formation
timescale. Studies of TTS have derived a disc dissipation timescale of
$\approx$ 5 Myrs \citep{2001ApJ...553L.153H}, with very few discs
shown to survive out to $\approx$ 20 Myrs
\citep{2004ApJ...612..496M}. However, the derivation of a disc
dissipation timescale from disc fractions suffers from several
intrinsic problems, for instance the inconsistent (and often
contradictory) methods of deriving ages \citep[see][for a full
discussion]{2008MNRAS.386..261M}. Notwithstanding this, disc fraction
surveys have also been published for lower mass and BD systems, such
as \cite{2004A&A...427..245S} and \cite{2004ApJ...609L..33M}, deriving
a similar order-of-magnitude disc dissipation timescale. The
probability of forming planets around BDs is clearly sensitive to any
changes in disc dissipation timescale across the TTS to BD mass
boundary.

Some studies of disc lifetimes find no significant difference between
BD stars and their higher mass counterparts
\citep{2003AJ....126.1515J}, however a possible correlation of disc
lifetime with stellar mass is proposed in \cite{2007ApJ...660.1517S}.
Additionally, \cite{2009arXiv0901.2603K} find that, for TTS older than
3 Myrs, the disc dispersal time is mass dependent for intermediate
mass (IM) to solar-type stars. This result only holds for the mass
transition stated (IM to solar-type) and is based on stars with
combined dust signatures, from infrared (IR) excesses and accretion
indicators (using a EW[H$\alpha$], criterion).
\cite{2009arXiv0901.4120D} also find, for TTS, a mass dependent
general SED shape for disc candidates, with higher mass stars showing
a featureless continuum at $\lambda>8.0$\,$\mu$m and lower mass stars
generally showing strong silicate and PAH features at 10 and 2\,$\mu$m,
indicative of mass dependent disc evolution. If the disc lifetimes and
therefore rate or type of involved physical processes are a function
of stellar mass. This could provide another piece of evidence
separating the previously assumed continuous (in mass) evolutionary
principles of stars. Resulting in a dichotomy separated by the
hydrogen-burning mass limit. This would also lead to a change in the
probability of forming planets in the BD regime, compared to CTTS.

To test the hypothesis that disc dissipation timescales are
independent of stellar mass one again requires accurate masses for BD
with associated discs (hereafter, BDD) systems, alongside reliable
disc fractions. The derivation of masses for BDD systems is
complicated by disc presence, but additionally, the detection of disc
presence at a given age is not simply a function of the disc
dissipation timescale.

Several mechanisms may act to completely dissipate a circumsteller
disc. These include accretion from the disc ($\delta M_{\rm
  disc}=\dot{M}$), photoevaporation or photoionisation due to flux
from the central and neighbouring stars, and the formation of giant
planets or interaction with binary partners. However, the entire disc
does not have to be dissipated to be missed by disc fraction surveys.
More subtly, disc fractions can be altered due to wavelength dependent
detection. This is due to the peak flux of the thermal blackbody
emission of the disc being proportional to the temperature,
$\lambda^{\rm thermal}_{\rm peak}\propto T_{\rm disc}$, and the
temperature in the disc being approximately a function of distance
from the star. Thus for a given system, as one increases the
wavelength of observation, the corresponding radial position of the
effective source will increase. Moreover, as the disc flux must be
separated from the photospheric flux the important quantity is the
separation in wavelength between the emission peaks for the stellar
and thermal disc components. Therefore the presence of an inner disc
hole can change derived disc fractions, as a function of
wavelength. Interestingly, \cite{2009arXiv0901.2603K} find that
results for disc dissipation timescales based only on their accretion
indicators or inferred disc presence do not agree, perhaps due to
inclusion of transitional objects, where the inner disc is being
cleared, and is therefore not detected at their target
wavelength. \cite{2009arXiv0901.2603K} suggest that this is caused by
photoevaporation of the inner disc, leading to a mass dependent disc
lifetime, at their wavelength, for their sample. However, as discussed
in \cite{2009arXiv0901.4120D}, the inner hole can be created via dust
grain growth and mid plane settling, giant planet formation or
photoevaporation by ultraviolet (UV) and far-UV (FUV) flux
\citep{2007MNRAS.378..369N}.  \cite{2009arXiv0901.4120D} argue that if
binarity or giant planet formation or photoevaporation cause inner
disc clearing the resulting system would not show active accretion
from radii below the inner disc boundary. For the case of clearing by
dust grain growth and settling accretion from a hot gas disc within
the inner radius is still possible, and indeed evidence for such a hot
gas inner disc has been found through SED fitting
\citep{2005ApJ...623..952E,2008ApJ...689..513T}. Additionally,
\cite{2009arXiv0901.4445E} argue that for M stars, with a pure
reprocessing disc where the inner disc radius ($R_{\rm inner}$) is
equal to the dust sublimation radius ($R_{\rm dust}$, discussed in
Section \ref{physics}), the stellar and thermal components are not
separable at wavelengths shorter than 6\,$\mu$m. In the case of
\cite{2009arXiv0901.4445E}, the inner disc hole was created by
photoionising radiation from the central star destroying the dust
component of the disc. Recent direct evidence for a photoevaporating
disc has been found for the circumstellar disc around a low mass TTS
in the Sigma Orionis cluster \citep{2009arXiv0902.0457R}. The
detection was characterised by a tenuous photoevaporation wind and
little or no accretion, and a disc detection at 24\,$\mu$m indicative of
a large inner hole. Additionally, for BDs, \cite{2009arXiv0902.0638M}
find that the disc mass and radius distributions in the Orion
Trapezium cluster are consistent with photoevaporation over 1 Myr (the
adopted age of the cluster), from UV and FUV flux generated by the
massive $\theta^1$ Ori C.

Whichever mechanism is responsible for clearing the inner disc hole it
is clear that its presence leads to confusion in the validity of disc
fractions at a given wavelength. The important quantity here is the
separation in wavelength between the emission peaks for the stellar
and thermal disc components, as it is these components which must be
separated. For BDD systems it is likely that difficulties detecting
the disc will be exacerbated by the lower temperatures of the
photosphere and therefore the smaller separation in wavelength from
the thermal disc emission component, compared to TTS
systems. Therefore, to derive disc fractions for BD populations we
require detailed comparison models with which to guide disc candidate
selection. These models must necessarily include a treatment of
accretion and dust destruction, given that these factors deplete the
inner (and total) disc thereby affecting disc detection.

Although the creation of the inner disc hole could have several
causes, at least part of the effect is undoubtedly due to
accretion. Accretion acts to directly deplete the inner disc (where
the material is accreted from). Accretion also depletes the inner disc
indirectly, as accretion generates hot continuum emission
\citep{1998ASPC..154.1709G}, increasing the photoionising UV and FUV
flux available \citep{1998ASPC..154.1709G}, and therefore affecting
the derived disc fraction \citep{2009arXiv0901.3684H}. This effect may
be amplified for BDD systems where the combined photospheric emission
is dominated by accretion flux. However, accretion columns modeled by
\cite{2004MNRAS.348..879A} were found to produce negligible levels of
photoionising flux due in part to the high optical depth of the column
itself to Lyman continuum photons. Additionally, due to the high
concentrations of hydrogen in the accretion column and hotspot the
emission is more likely to be characterised by a stellar atmosphere
than a black body. However, for our work as emission from a stellar
atmosphere or a black body of equal temperature are approximately
equal for wavelengths longer than the Lyman edge we have neglected
this effect. Whilst the accretion column and hotspot may not
contribute to the photoinising flux within the system they will of
course alter the disc evolution by an increase of the emitted
flux. This increase in flux will lead to an increase in temperature of
the disc inner and outer regions, and more particularly a change in the
dust sublimation effects at the inner edge.

The current proposed mechanism of accretion is a magnetically
channelled accretion `column', where material from a truncated (at the
co-rotation radius $R_{\rm co}$) inner disc boundary falls onto the
surface of the star along magnetic field lines
\citep{1990RvMA....3..234C,1991ApJ...370L..39K}. This mechanism is
widely accepted, with strong evidence found from the excessive widths
of H$\alpha$ ($>$10\,\AA) emission lines, over that caused by
chromospheric activity and stellar rotation
\citep{1998ASPC..154.1793W}, caused by the high velocities of
infalling matter, $~$100-200 kms$^{-1}$ \citep{1998ApJ...492..743M}.

Accretion rates can be derived by modelling this magnetospheric
accretion with a characteristic `column' temperature and assuming
statistical equilibrium of the state populations. The resulting line
ratios can then be fitted to the observed spectra. Accretion rates can
also be calculated from model comparison of the Equivalent Width (EW)
of the H$\alpha$ line itself, or perhaps more reliably, as they are
less affected by extinction, that of the hydrogen recombination lines
of Pa$\beta$ and Br$\gamma$. Accretion rates have been derived using
these methods for many TTS and more recently for BDs
\citep{2003ApJ...592..266M,2004A&A...424..603N,2006A&A...452..245N},
with accretion rates in the range $\dot{M}_{\rm
    acc}=$10$^{-6}$--10$^{-12} M_{\odot}yr^{-1}$, and typical
accretion rates in the range \logmdot = log$(\frac{\dot{M}}{M_{\odot}}
yr^{-1})=-10$ to $-8$ with only FU Orionis type objects accreting at
rates higher than this. The mechanism producing accretion rates as
high as those observed episodically in FU Orionis type objects is
likely to be produced by some accretion mechanism instability, or
perhaps by encounters with other cluster members \citep[see][for
discussion]{2008arXiv0810.2854P}. Additionally, much lower accretion
rates of $\dot{M}\approx$ 10$^{-13}M_{\odot}yr^{-1}$ have been derived
for BDs using flux blueward of 4000\,\AA where the photospheric flux
becomes negligible \citep{2009arXiv0901.3684H}.

Analysis of the derived accretion rate data has led several authors to
propose a strong correlation between accretion rate and stellar mass
\citep{2003ApJ...592..266M,2004A&A...424..603N,2006A&A...452..245N},
of an approximate form: $\dot{M_{\rm acc}}\propto M_{*}^{~2}$. This
correlation has no clear physical basis and it has been suggested that
it could be caused solely by selection biases and detection thresholds
\citep{2006MNRAS.370L..10C}. \cite{2006ApJ...648..484H} points out
that Bondi-Hoyle accretion naturally results in a clear
$\dot{M}\propto M^2_*$ relationship, but is not a desirable formalism
to adopt as it completely ignores angular momentum of the accreting
material. This proposed correlation could have important implications
with regards to disc lifetimes and planet formation probabilities as a
function of mass.

As discussed, discs associated with BD systems lead to difficulties in
accurate classification of the underlying BD
\citep{2004MNRAS.351..607W}. Furthermore, accretion, both direct and
indirect, can act to alter the disc signature, and clearly the
photospheric signature. For cooler BD systems, where the stellar flux
is generally redward of any accretion flux, increased accretion fluxes
could dominate the resulting SED. Given that the derivation of a
robust relationship of the form $\dot{M}\propto M^{\alpha}_*$ requires
accurate mass and accretion derivations, then clearly the effects of
circumstellar disc and increased accretion rates must be considered,
before any such relationship can be accepted. As there is currently no
clear theory linking accretion rate and mass (ignoring Bondi-Hoyle
accretion), this relationship is driven by observations.  However, the
observational constraints must be interpreted using a self-consistent
set of accreting BDD and BD systems to better understand their
implications. In other words, what do the observations really tell us?
Is there a real relationship between mass and accretion rate or would
simulated systems outside this relationship (and attributed scatter)
not be detected or included in the current observational surveys, due
to misclassification?

A further use of disc fractions, in conjunction with stellar ages, is
to aid the understanding of the angular momentum evolution of star and
disc systems. Recent evidence from TTS, from correlations of stars
with IR excesses and slower rotational rates
\citep{2007ApJ...671..605C} has seemed to strongly support a mechanism
of disc-locking, where the central star is rotationally braked through
magnetic locking to the accretion disc
\citep{1993AJ....106..372E,1990RvMA....3..234C,1991ApJ...370L..39K}. A
similar mechanism has been put forward, but has yet to be
unambiguously confirmed, for BD stars. However, current studies have
found a much weaker correlation between the slower rotators and those
where a disc presence is inferred for BDs. Some authors have therefore
proposed an `imperfect' disc-locking mechanism for stars in the lower
mass regime ($M_*<0.4M_{\odot}$, inclusive of the BD mass regime,
$M_*<0.08M_{\odot}$) \citep{2005A&A...430.1005L}. This weaker
correlation could also be due to `imperfect' disc-locking or
observational effects and mass dependence of the disc dissipation
timescale \citep{2004A&A...419..249S,2005A&A...429.1007S}.

If disc-locking is present for BDD systems, as for TTS stars, one
would expect accretion of material from a magnetic truncation radius,
$R_{\rm T}$ \citep{2007prpl.conf..479B}. \cite{2007prpl.conf..479B}
show that the magnetic truncation radius, $R_{\rm T}$, must be less
than the co-rotation radius, $R_{\rm co}$ (where the Keplerian angular
rotation rate of the disc is the same as the angular rotation rate of
the star), as material outside this radius will have too much angular
momentum to fall onto the star. This is implicitly adopted in the
proposed accretion mechanism, where the disc is magnetically
truncated, and therefore locked to the star. This magnetic truncation
would lead to an inner wall (and inner hole) of the accretion disc.
Evidence for an inner disc wall has been found from near-IR excesses
from the inner rims of Herbig Ae stars, with emission peaking at
2--3\,$\mu$m \citep{2001ApJ...560..957D}. The actual shape and radial
position of this wall is affected by several factors, discussed
further in Sections \ref{dust_edge} and \ref{inner_edge}. The previous
work of \cite{1997AJ....114..288M} explored the effects of simple mass
accretion and the radius to the inner disc on the infrared colours of
CTTS, using a prescribed range of inner disc radii.
\cite{1997AJ....114..288M} find that the scatter in infrared colours
for there parameter range, $R_{\rm inner}=$1$-$12$R_*$ and
$\dot{M_{\rm acc}}=$10$^{-9}$to 10$^{-5} M_{\odot}yr^{-1}$, matches
the scatter of the CTTS locus found in the data. The models of
\cite{1997AJ....114..288M} used physically flat discs however meaning
their results will not stand for discs with significant flaring. More
importantly for our work \cite{1997AJ....114..288M} find a correlation
of the infrared locus of models with the accretion and inner disc
location. Essentially, changes in the stellar flux levels and position
of the inner wall lead to temperature changes of this inner edge and
therefore changes in the balance of flux across the infrared
wavelength regime. As this study does not include the effects of
vertical hydrostatic equilibrium or even the study of flared discs one
would expect, perhaps, a more complex relationship between the inner
edge position and infrared excess in `real' systems. This means that
studies aiming to provide evidence of disc locking by comparing disc
indicators with rotation rates may include unforeseen biases. As the
rotation rate will effect the co-rotation radius and therefore
truncate the disc at different radii, thereby enhancing or suppressing
any relationship of infrared excess (or disc presence) with rotation
rate.

In summary, several key areas in the theoretical studies of BDs hinge
critically on observational constraints, where the derived values are
heavily affected by disc and/or accretion presence. Formation
mechanisms, planet formation, IMFs, disc dissipation timescales,
accretion rate--stellar mass relationships and studies of disc-locking
mechanisms using correlations of rotation rates with infrared
excesses, all require derivations of either masses, ages or disc
fractions (or a combination), supplemented by other parameters. The
consequences of the disc structure for BD accretion discs,
particularly vertical structure, given the assumption of vertical
hydrostatic equilibrium, have already been discussed for a limited
range of BD parameters in \cite{2004MNRAS.351..607W}. Indeed
\cite{2004MNRAS.351..607W} have shown that BDD systems seen at higher
inclinations can be confused with CTTS systems, having similar colours
and magnitudes. A benchmark model grid of spectra for star and disc
systems in the mass range 0.1$M_{\odot}<M_*<$50$M_{\odot}$ has
previously been published in \cite{2006ApJS..167..256R}.

In this work we explore the effects of accretion and disc presence on
the derivation of the stellar parameters and explore the intrinsic
biases, or relationships, which will effect or inhibit conclusions on
mass to accretion rate or rotation to infrared excess correlations. We
have extended the similar work of \cite{2004MNRAS.351..607W} by
inclusion of a simple treatment of accretion and a sophisticated
implementation of dust sublimation (amongst other improvements). The
complete model grid, with derived photometry and isochrones, is
available online through our browsing tool
\footnote{http://www.astro.ex.ac.uk/research/bd\textunderscore disc}
which is described in Appendix \ref{website}. The aim of this online
database is to complement that of \cite{2006ApJS..167..256R}. Finally,
subsequent papers will introduce expansion to the model grid and an
online fitting tool.

**THIS PARAGRAPH IS CURRENTLY WRONG, WILL REQRITE AT THE END**

The structure of this paper is as follows. In Section \ref{model} we
outline the model used, including an explanation of our physical
assumptions (Section \ref{physics}), a brief description of the
radiative transfer code (Section \ref{torus}), and an explanation of
the derivation of emergent quantities such as photometric magnitudes
(Section \ref{derived}). The ranges of the input variables are then
explained and, where possible, justified using observational
constraints in Section \ref{par_space}, with the total parameter space
covered shown in Table \ref{par_space_table}. Section
\ref{disc_struct} then contains analysis and description of the
resulting temperature and dust density structures of the discs. The
consequences of these changes in disc structure on the SEDs of these
objects is then discussed in Section \ref{sed_analysis}. Section
\ref{photometry} then details problems caused for the derivation of
the primary parameters age, mass and disc fractions from photometry
and the resulting relations built on these parameters; accretion rate
to stellar mass and rotational period to disc presence. Finally,
conclusions are drawn, with a summary, in Section \ref{conclusions}..

@@

In this Section we detail the physical model adopted and assumptions
made (Section \ref{physics}), then explain key elements of the
radiative transfer code (Section \ref{torus}). Then we discuss the
derived values, such as broadband photometric magnitudes and colours
(in Section \ref{derived}). Some internal consistency checks are
detail in Appendix \ref{consistency}.

@@

To model an accreting BDD system we must first model the underlying
photospheric flux. To provide a self-consistent set of model spectra
we have adopted a BD stellar interior and atmospheric model grid. We
have then constructed the total photospheric flux for any input value
of stellar age and mass by interpolating for surface gravity (log(g)),
effective temperature ($T_{\rm eff}$), radius ($R_*/R_{\odot}$) and
luminosity ($L_*/L_{\odot}$). These values were then used to
interpolate atmospheric spectra for flux (ergs s$^{-1}$cm$^{-2}
{\rm \AA}^{-1}$) from 1200 to 2$\times$ 10$^7$ \AA (or 0.12 to 2000\,
$\mu$m). These spectra were then smoothed using 200 logarithmically
spaced points. Careful inspection ensured that no spectral features
were removed during smoothing. The stellar interior models used for
this study are the `DUSTY00' models of \cite{2000ApJ...542..464C}
combined with the `AMES-Dusty', atmospheric models of
\cite{2000ApJ...542..464C}, which are all available online
\footnote{http://perso.ens-lyon.fr/france.allard/}. For our
approximate $T_{\rm eff}$ (effective temperature) range of $\approx
$3000\,K$<T_{\rm eff}<$1600\,K the AMES-Dusty atmospheres are most
applicable (2700\,K$>T_{\rm eff}>$1700\,K). We did try including dynamic
application of atmospheres based on the derived $T_{\rm eff}$,
i.e. using AMES-Cond for $T_{\rm eff}<$1700\,K, however this resulted in
large discontinuities between the model atmospheres and resulting
spectra. As this only affects stars at the very edge of our
temperature range, i.e. for the oldest and lowest mass objects (for
the AMES-Cond case), we have adopted the same model throughout.

@@

Secondly, we require a model of the accretion flux for the system. The
flux due to accretion was added onto the intrinsic photospheric flux
using the simple approximation of blackbody emission. The selected
accretion rate was used to derive an accretion luminosity ($L_{\rm acc}$),
where the material was modelled as falling from the disc inner edge
onto the surface of the star. $L_{\rm acc}$ is calculated according to
\ref{Lacc}, 
\begin{equation}
{L_{\rm acc}=\frac{GM_*\dot{M}}{R_*} \left( 1-\frac{R_*}{R_{\rm inner}}\right)},
\label{Lacc}
\end{equation}
where $M_*$ is the stellar mass, $\dot{M}$ the mass accretion rate,
$R_*$ the stellar radius and $R_{\rm inner}$ the radius of the disc inner
boundary.

The initial inner disc radius was set to be the co-rotation radius
(this is discussed in more detail in Section
\ref{disc_parameters}). During the radiative transfer simulations of
the disc the final inner dust disc radius may be extended past the
co-rotation radius. This is due to dust sublimation effects (see
Sections \ref{dust_edge} and \ref{inner_edge} for an
explanation). However, as the gas within the disc is not sublimated it
will always extend down to the co-rotation radius and we assume
accretion will still occur from this radius. A full discussion of the
differences between the initial and resulting final radii of the dust
disc can be found in Section \ref{inner_edge}. Once the accretion
luminosity was derived, an adopted areal coverage ($A$), over the
stellar surface, was used to calculate an effective temperature
($T_{\rm acc}$), for the accretion `hot' spot, where
\begin{equation}
{T_{\rm acc}=\left(\frac{L_{\rm acc}}{4\pi R^2_{*} \sigma A}\right)^{\frac{1}{4}}}.
\label{Tacc}
\end{equation}

Finally, a blackbody flux distribution is generated at $T_{\rm acc}$
and added onto the intrinsic stellar photospheric flux. In general one
would expect this to be an overestimate of the accretion flux, as for
pre-MS BDs large convective zones are expected on the stellar surface,
and some of the accretion energy may act to further drive these
convective currents, meaning flux is lost. It is worth noting however
that observationally UV excesses are often used to recreate and then
subtract an assumed accretion flux using a blackbody flux curve,
essentially the reverse of this method.

@@

We have used the {\sc torus} radiative transfer code which is
described in \cite{2000MNRAS.315..722H}, including the subsequent
refinements i.e.  addition of Adaptive-Mesh-Refinement (AMR)
introduced in \cite{2004MNRAS.350..565H}. {\sc torus} uses the
criterion of \cite{1999A&A...344..282L} in a radiative equilibrium
algorithm. The simulation also calculates \textit{vertical hydrostatic
  equilibrium} and dust sublimation for the disc, similar to that
described in \cite{2007ApJ...661..374T}.

@@

\subsubsection{Dust sublimation and the Inner Disc Edge}
\label{dust_edge}

We have also applied dust sublimation during our simulations. This
treatment is similar to that detailed in \cite{2007ApJ...661..374T},
but improved and generalised with respect to the applicable
geometries.

>>>>

\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2010}

@@

There is now strong evidence that as brown dwarfs (BDs) form they pass
through a Classical T~Tauri star (CTTS) phase, during which they
possess a flared, dusty circumstellar disc from which they are
actively accreting material (xxx; xxx). The accretion is thought to
proceed via a magnetically-controlled funnel flow mechanism in which
material from a truncated inner disc boundary falls onto the surface
of the star along magnetic field lines
\citep{1990RvMA....3..234C,1991ApJ...370L..39K} xxx.

The truncated inner-edge of the circumstellar disc receives direct
photospheric radiation and is strongly heated. This leads to an
increased scaleheight, or `puffing up' of the inner disc rim, and
subsequent shadowing of the region immediately beyond it (Dullemond
reference). The density-dependent nature of the dust sublimation
temperature (pollack ref) may also play a role in shaping the inner
rim (isella ref; tannirkulam ref).

As an aside, although we recognise that the inner hole in protostellar
discs may be the result of giant planet formation or photoevaporation
(xxx; xxx), these two mechanisms should result in a negligible
mass-accretion rate and therefore would represent a distinct
evolutionary state of brown dwarf disc (BDD) sytems to that of the
CTTS-like phase we are considering here.

The interpretation of the spectral energy distributions (SEDs) of
pre-main-sequence (PMS) stars involves trying to distinguish the
various contributions to the continuum from the hot spots at the base
of the accretion flow, the photospheric flux, and the near-IR flux
from the dusty inner disc. The complexity of the interplay between
these contributions is excerbated by other, geometrical, effects such
as the inclination (which changes the projected area of the inner disc
wall visible to the observer) and the outer disc structure (which may
obsure the inner disc for high inclincations). (xxx; xxx)

In surveys, particularly those attempting to discern the PMS disc
fraction (Xxx), this disentangling usually takes place using broad-band
photometric measures and cuts in colour-colour or colour-magnitude
space (xxx; xxx).  The important quantity here is the separation in
wavelength between the emission peaks for the stellar and thermal disc
components, as it is these components which must be separated. For BDD
systems it is likely that difficulties detecting the disc will be
exacerbated by the lower temperatures of the photosphere and therefore
the smaller separation in wavelength from the thermal disc emission
component, compared to CTTS systems. Therefore, to derive disc
fractions for BD populations we require detailed comparison models
with which to guide disc candidate selection.

Comparison of accretion rates across the pre-main-sequence (PMS) mass
spectrum has indicated that the accretion rate is strongly correlated
with PMS mass,
\citep{2003ApJ...592..266M,2004A&A...424..603N,2006A&A...452..245N},
with an approximate form $\dot{M_{\rm acc}}\propto M_{*}^{~2}$. Since
in the canonical picture the accretion rate is driven by the disc
viscosity, and should be indepedent of the mass of the central object,
this correlation is somewhat surprising. There is some danger that the
correlation is the result of, or at least strengthened by, the
presence of observational biases (xxxx): At the high-mass (CTTS) end of the
mass spectrum the lowest accretion-rates cannot be measured via
continuum methods since the excess is too small in contrast with the
photopsheric emission. The emission lines will also be weak, and
indeed the H$\alpha$ equivalent width (EW) may be less than the
10\AA\ that traditionally demarc?ates classical from weak-lined T Tauri
stars. Such objects should show doppler-broadened profiles, but the
line wings will be weak, and the presence of underlying H$\alpha$
absorption may become dominant. Thus there may be a population of
low-accretion-rate CTTS which are current missing from the surveys.

There are also risks of observational biases at the low-mass end of
the correlation. Radiative-transfer modelling of H$\alpha$ emission
from accreting BDs requires high temperatures ($>$\, 10\,kK) in the
accretion funnels in order to recover the level of emission that is
observed. (Cooling rate arguments can be invoked to explain the
presence of such high temperatures). Recently some continuum
measurements of accreting BDs have been conducted, and these broadly
support the H$\alpha$ rates (xxxx). Nonetheless it is worth considering the
observed effect that mass accretion at typical CTTS rates onto a BD
might have. Parity between the BD photospheric luminosity and the
accretion luminosity occurs at relatively low accretion rates
\citep{2006MNRAS.370L..10C}. Such objects may not be detected as BDs
at all in photometric surveys. Even before such an extreme case is
reached, the additional luminosity provided by the accretion should
have measurable effect on the BD colours. Furthermore the additional
flux will be reprocessed by the disc, altering its scaleheight and
possibly the shape of its inner edge.

Here we present a grid of models of BDD systems, including a
self-consistent treatment of the photospheric and accretion luminosity
sources and the interaction of that flux with the circumstellar disc.
We investigate the impact of the luminosity of the central source on
location and shape of the disc inner rim, as well as the large-scale
structure and flaring of the outer disc. We construct synthetic
colour-colour and colour-magnitude diagrams in order to examine the
efficacy of the photometric selections used to isolate brown dwarfs
and measure disc fractions (xxx;xxx).

The complete model grid, with derived photometry and isochrones, is
available online through our browsing tool
\footnote{http://www.astro.ex.ac.uk/research/bd\textunderscore disc}
which is described in Appendix \ref{website}.

@@

In this Section we detail the physical model adopted and assumptions
made (Section \ref{physics}), then explain key elements of the
radiative transfer code (Section \ref{torus}). Then we discuss the
derived values, such as broadband photometric magnitudes and colours
(in Section \ref{derived}). Some internal consistency checks are
given in Appendix \ref{consistency}.

@@

In order to model an accreting BDD system we must first model the
underlying photospheric flux. We have adopted a BD stellar interior
and atmospheric model grid and have then constructed the total
photospheric flux for any input value of stellar age and mass by
interpolating for surface gravity ($\log g$), effective temperature
($T_{\rm eff}$), radius ($R_*/R_{\odot}$) and luminosity
($L_*/L_{\odot}$). These values were then used to interpolate
atmospheric spectra for flux (ergs s$^{-1}$cm$^{-2} {\rm \AA}^{-1}$)
from 1200 to 2$\times$ 10$^7$ \AA. The spectra were subsequently
resampled onto 200 logarithmically spaced points. Careful inspection
ensured that no spectral features were removed during resampling. The
stellar interior models used for this study are the `DUSTY00' models
of \cite{2000ApJ...542..464C} combined with the `AMES-Dusty',
atmospheric models of \cite{2000ApJ...542..464C}, which are all
available online
\footnote{http://perso.ens-lyon.fr/france.allard/}. For our $T_{\rm
  eff}$ range of $\approx $3000\,K$<T_{\rm eff}<$1600\,K the
AMES-Dusty atmospheres are the most applicable (2700\,K$>T_{\rm
  eff}>$1700\,K). We did try including dynamic application of
atmospheres based on the derived $T_{\rm eff}$, i.e. using AMES-Cond
for $T_{\rm eff}<$1700\,K, but this resulted in large discontinuities
between the model atmospheres and resulting spectra. Since this only
affects stars at the very edge of our temperature range, i.e. for the
oldest and lowest mass objects (for the AMES-Cond case), we have
adopted the AMES-Dusty models throughout.

@@

We assumed blackbody emission for the accretion flux.  The selected
accretion rate was used to derive an accretion luminosity ($L_{\rm
  acc}$), where the material was modelled as free-falling from the
disc inner edge onto the surface of the star. $L_{\rm acc}$ is
calculated according to \ref{Lacc},
\begin{equation}
{L_{\rm acc}=\frac{GM_*\dot{M}}{R_*} \left( 1-\frac{R_*}{R_{\rm inner}}\right)},
\label{Lacc}
\end{equation}
where $M_*$ is the stellar mass, $\dot{M}$ the mass accretion rate,
$R_*$ the stellar radius and $R_{\rm inner}$ the radius of the disc inner
boundary.

The initial inner disc radius was set to be the co-rotation radius
(this is discussed in more detail in Section
\ref{disc_parameters}). During the radiative transfer simulations of
the disc the final inner dust-disc radius may be beyond the
co-rotation radius due to dust sublimation effects (see Sections
\ref{dust_edge} and \ref{inner_edge} for an explanation). Once the
accretion luminosity was derived, an adopted areal coverage ($A$),
over the stellar surface, was used to calculate an effective
temperature ($T_{\rm acc}$), for the accretion `hot' spot, where
\begin{equation}
{T_{\rm acc}=\left(\frac{L_{\rm acc}}{4\pi R^2_{*} \sigma A}\right)^{\frac{1}{4}}}.
\label{Tacc}
\end{equation}

Finally, a blackbody flux distribution is generated at $T_{\rm acc}$
and added onto the intrinsic stellar photospheric flux. In general one
would expect this to be an overestimate of the accretion flux, as for
pre-MS BDs large convective zones are expected on the stellar surface,
and some of the accretion energy may act to further drive these
convective currents, meaning flux is lost. It is worth noting however
that observationally UV excesses are often used to recreate and then
subtract an assumed accretion flux using a blackbody flux curve, which
is essentially the reverse of this method.

@@

We have used the {\sc torus} radiative transfer code which is
described in \cite{2000MNRAS.315..722H}, including the subsequent
refinements i.e.  addition of Adaptive-Mesh-Refinement (AMR)
introduced in \cite{2004MNRAS.350..565H}. {\sc torus} uses the method
of \cite{1999A&A...344..282L} solve radiative equilibrium. The
simulation also self-consistently solves the equation of
\textit{vertical hydrostatic equilibrium} and dust sublimation for the
disc (described in \cite{2007ApJ...661..374T}).

@@

\subsubsection{Dust sublimation and the inner disc edge}
\label{dust_edge}

We have also applied dust sublimation during our simulations. This
treatment is similar to that detailed in \cite{2007ApJ...661..374T},
but with some enhancements to ensure convergence of the sublimated
rim.
